### 2023-10-19 conclusions/further.tex
Renamed files and folder

The final circuit will need an external controller to manage diffenrent things in the circuit. This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated %TODO ref annex

@@

Training directly on the chip has a great advange over external training. Indeed training within the chip will allow to completely ignore the wires own resistance in the crossbar array. This is because they will an integral part of the weights, if the weight needs to be lowered then the resistance will be raised (because the conductances need to be lowered). %TODO finish with better explainnation

>>>>

The final circuit will need an external controller to manage diffenrent things in the circuit. This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated.

@@

### 2023-10-19 design/dataset.tex
Renamed files and folder

\subsubsection{Results} %TODO Move to results obviously

The weights are trained using both the custom activation function (\cref{sec:genwei}) and without ones and compare their output.

After training on a digital computer using the settings described in \cref{sec:genwei}.

>>>>

\subsubsection{Timing}

The timing of the flags are quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.

\Cref{tim:airline} shows the time diagram of the entire execution using the parameters discussed earlier.

\begin{figure}[H]
  \centering
  \begin{tikztimingtable}%TODO change or find other way
    $e_0$       & x 2H 6L     L 2H 6L     L L x\\
    $e_1$       & x 2L 2H 4L  L 2L 2H 4L  L L x\\
    $e_2$       & x 4L 2H 2L  L 4L 2H 2L  L L x\\
    $e_3$       & x 6L 2H     L 6L 2H     L L x\\
    $e_{next}$  & x 8L        H 8L        H L x\\
    $e_{in}$    & x 8H        L 8H        L L x\\
    $e_{out}$   & x 8L        L 8L        L H x\\
    \extracode
    \tablerules
    \begin{pgfonlayer}{background}
      %\vertlines[help lines]{0.55,10.55}
      %\vertlines[red]{1.6,5.6,15.6}
      %\vertlines[blue]{3.6,9.6,15.6}
    \end{pgfonlayer}
  \end{tikztimingtable}
  \caption{Flags time diagram for the airline problem with $n_s=4$}
  \label{tim:airline}
\end{figure}

### 2023-10-19 design/netlist.tex
Renamed files and folder

This is done by generating a SPICE netlist using a python script. This script takes in a few parameters :

\begin{itemize}
  \item The first parameter is the number of input we use our system. This is the size of the first input vector ($x_t$).
  \item The next parameter is the number of time steps. Everything about the \ac{LSTM} time steps is all explained in \cref{sec:lstm}.
  \item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network. This parameter must divide the number of hidden states in the \acp{LSTM} layers.
  \item The files containing different info about the model. This file contains the type of \ac{NN} architecture and the weights associated with each layer. The weights in the files have to be organized using a specific model (\cref{subsec:weiStore}).
  \item Finally the name of the file in which the output of the script (the netlist) will be written to.
\end{itemize}

>>>>

This is done by generating a SPICE netlist using a python script. A SPICE netlist was chosen because of Cadence's virtuoso limitations. Indeed, Verilog netlists can be imported with a downside, component's parameters can't be imported. This is very limiting for this thesis' use case (required to set the resistors' resistance and thus the weights). SPICE netlists can import parameters for the components and is open source and by such well documented. The generator script takes in a few parameters :

\begin{itemize}
  \item The first parameter is the number of input we use our system. This is the size of the first input vector ($x_t$).
  \item The next parameter is the number of time steps. Everything about the \ac{LSTM} time steps is all explained in \cref{sec:lstm}.
  \item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network. This parameter must divide the number of hidden states in the \acp{LSTM} layers.
  \item The files containing different informations about the model. This file contains the type of \ac{NN} architecture and the weights associated with each layer. The weights in the files have to be organized using a specific model (\cref{subsec:weiStore}).
  \item Finally the name of the file in which the output of the script (the netlist) will be written to.
\end{itemize}

\subsection{Available layers}

As of the writing of the thesis, the netlist generator can generate the netlist for :

\begin{itemize}
  \item \ac{NP} \ac{LSTM}
  \item Vanilla \ac{LSTM}
  \item \ac{FGR} \ac{LSTM}
  \item \ac{GRU}
\end{itemize}

Nevertheless, the weights can only be generated for the \ac{NP} \ac{LSTM} and the \ac{GRU} (\cref{sec:genwei}). While the script can tecnically generate the Vanilla and \ac{FGR} \acp{LSTM}, since there is no way to generate weights for those version of the \ac{LSTM}, they are rendered useless.

### 2023-10-19 design/wei2res.tex
Renamed files and folder

For the thesis, this step is done in python and integrated in the netlist generator script (\cref{sec:netlist}). Since the resistance resolution is of about%TODO check memristor precision
, the python code then reproduces this by limiting the amount of significant digit of the resistance.

>>>>

For the thesis, this step is done in python and integrated in the netlist generator script (\cref{sec:netlist}). The resolution of the memristor (the precision of the resistances at which the memristor can be set) is simulated by limiting the generated resistances values to having only two significant figures.

### 2023-10-19 results/airline.tex
Renamed files and folder

>>>>

coincidentally, the curve is closer to the target curve than the digital prediction itself, this is mostly due to the lack of precision of the analog prediction, but it is nevertheless interresting to mention. \Cref{graph:airlineCoin} shows the graph and the \ac{RMSE} is the lowest, even lower than the ones from the digital predictions.

\begin{figure}[H]
  \centering
  \includesvg[width=\linewidth]{results/airline/closeRes}
  \caption{Analog prediction ($n_s=1$) with the target dataset.}
  \label{graph:airlineCoin}
\end{figure}

### 2023-10-19 state/gru.tex
Renamed files and folder

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
  \label{fig:lstmCell}
  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

@@

The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}. \cite{gru,gruKeras,gruPyTorch} As it's name implies, \acl{CIFG}, links the input and forget gates as such, $f_t=1-i_t$. Other differences are :
\begin{itemize}
  \item No peepholes and output activation function.
  \item Combined hidden state and cell state.
  \item The candidate cell state is point wise multiplied with the output gate before the activation function.
\end{itemize}

This is clearer with the math equations :
\begin{equation}\label{eq:cellGGRU}
  \tilde{c}_t = tanh(w_{cx}\cdot x_t + w_{ch}\cdot(o_t \odot h_{t-1}) + b_c)
\end{equation}
\begin{equation}\label{eq:hiddenGRU}
  h_t=(1-i_t)\cdot h_{t-1} + i_t \cdot \tilde{c}_t
\end{equation}

>>>>

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
  \label{fig:lstmCell}
  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

@@

While the \ac{GRU} is its own type of \ac{RNN}, it has been compared and assimilated to an \ac{LSTM} \cite{nbLSTM}. Indeed, the subtype of \ac{LSTM} it has been associated with is called the \ac{CIFG} \ac{LSTM}. If the input and forget gates are combined into one ($f_t=\overrightarrow{1}-i_t$), they can be assimilated as the update gate of the \ac{GRU}. The reset gate would then be the \ac{LSTM}'s output gate.

\Cref{eq:lstmGru0,eq:lstmGru1,eq:lstmGru2,eq:lstmGru3,eq:lstmGru4} are the \ac{GRU}'s equations with the \ac{LSTM}'s names.

\begin{equation}\label{eq:lstmGru0}
  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_f + \overrightarrow{b_f})
\end{equation}
\begin{equation}\label{eq:lstmGru1}
  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_o + \overrightarrow{b_o})
\end{equation}
\begin{equation}\label{eq:lstmGru2}
  \overrightarrow{i_t}=\overrightarrow{1}-\overrightarrow{f_t}
\end{equation}
\begin{equation}\label{eq:lstmGru3}
  \overrightarrow{\tilde{c_t}}=tanh(\overrightarrow{x_t}\cdot w_{cx}+(\overrightarrow{o_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{ch} + \overrightarrow{b_c})
\end{equation}
\begin{equation}\label{eq:lstmGru4}
  \overrightarrow{h_t}=(\overrightarrow{i_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{f_t}\odot \overrightarrow{\tilde{h_t}}
\end{equation}

Differences between the two \acp{RNN} are :
\begin{itemize}
  \item The \ac{GRU} has no output activation function.
  \item The \ac{GRU} has combined hidden state and cell state.
  \item The \ac{GRU}'s candidate hidden state is point wise multiplied with the reset gate before passing into the dense layer.
\end{itemize}

### 2023-10-19 state/model.tex
Renamed files and folder

The voltage dependence of the memristor's model is described in \cref{graph:voltChange}. The influence of the bias voltage is, as \cref{eq:memModel3, eq:memModel4} show, the way to change the resistance range of the memristor, until the physical limit is reached of course. Applying higher voltage also allows to reach a set resistance faster as the slopes are bigger the higher the bias voltage gets.

>>>>

The voltage dependence of the memristor's model is described in \cref{graph:voltChange}. The influence of the bias voltage is, as \cref{eq:memModel3,eq:memModel4} show, the way to change the resistance range of the memristor, until the physical limit is reached of course. Applying higher voltage also allows to reach a set resistance faster as the slopes are bigger the higher the bias voltage gets.

The resistive state can also be lowered by applying negative voltages.TODO

### 2023-10-19 state/nn.tex
Renamed files and folder

The weights values are obtained through training. This is done in a few steps :

\begin{enumerate}
  \item Generate random values for each weight.
  \item Run the \ac{NN} to get an output vector.\label{step:restart}
  \item Measure error rate using the chosen error algorithm.
  \item Use the
  \item go back to \cref{step:restart}
\end{enumerate}

>>>>

The weights values are obtained through training. To train the weights it is required to have a dataset of inputs/outputs values that are correct. The outputs are the target values, the values that the \ac{NN} need to compute when being executed. Once the weights have been initialized (usually randomly picked values). The training happens by running the following steps (known as an epoch) :

\begin{enumerate}
  \item Run the \ac{NN} to get an output vector.\label{step:restart}
  \item Measure error rate using the chosen error algorithm.
  \item Run the backpropagation algorithm to determine how to change each individual weight.
\end{enumerate}

A \ac{NN} is trained for at least hundreds of epochs to get decent results.

Once trained, the \ac{NN} can be tested by feeding the \ac{NN} input data it hasn't been trained with.

One layer of this feedforward \ac{NN} is also known as a dense layer.
